\bf Adaption of the Chumbley Score to matching of bullet striation marks

### 2018-02-05 doc/paper.Rmd
adding modified introduction

## Motivation

Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results.

In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm. The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others). Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive.

Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm. Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes.

An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features. The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks. \citet{aoas}

The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region. The markings or striae in this region are supposed to have a high CCF with each other. Therefore a Bullet profile is chosen from this region by taking the cross section at a given height. A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm.

Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not. The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source. \citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm.

\citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
```{r}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)

knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold"))
```

@@

```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
  #gather(type, error, actual:beta) %>% 
  gather(type, error, beta) %>% 
  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
    theme_bw()+ 
  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")

@@

combined120
#wo120_profile
#wo120_1
#multiplot(wo120_1, wo120_profile1)
```
```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}

miss_1
```
```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}

miss_2
```
```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}

>>>>

## Motivation
Compairing pairs of toolmarks with the intention of matching it to a tool has been studied many times in the past. Extensive examples in literature for tools and toolmark research ranging from screwdrivers to groove pliers to slip-joint pliers can be found in the work of  \citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller}, \cite{afte-chumbley} and many more. In comparison to this, same source matching of bullets to firearms has not been examined as prominently as that of toolmarks, with even less information available on validity of methods and error rates associated with firearms examination. The National Academy of Sciences in its report in 2009 \citep{NAS:2009} discussed the need for determining error rates in methods proposed for firearms examination. In the context of same source matching and error rates determination, as seen in the case of most forensic applications, the first step involves identification of unique features that are characteristic of the object at hand. For the case of bullets and firearms, striation marks on the surface of the bullet are considered to be such markings that can be used in methods for same source matching. These marks are often a product of rifling and impurities and defects due to manufacturing in the barrel of the gun, which leads to engravings on the bullet surface \citep{afte-article1992}. In current practice, firearm examiners invariably make visual comparisons of bullet striae and use visual assesment tools to dignify bullets as being matches and non-matches. One way of accomplishing anykind of comparison between bullets is to do a comparison between surface marking of two or more bullet lands. Bullet Lands are considered to be areas between grooves made by the rifling action of the barrel. These marking are considered to be unique. The land engraved markings or sometimes termed as Bullet profiles \citep{aoas} are striation marks made on Bullet lands and often used for these land to land comparisons. Bullet Signatures is another word used in literature as seen in the work of \citet{chu2013} and \citet{aoas}. In our context bullet signatures refer to a processed version of the raw land engraved markings or profiles. The generation of bullet signatures involves first extraction of a bullet profile by taking the cross-sectional of the surface at a given height and then using loess fits to model the structure. The residuals of this fit are called signatures, which are considered to be noise free and a good reflection of the class charecteristics and unique features of a bullet. A more detailed version of the extraction technique of signatures is discussed by \citet{aoas}, where comprehensive details about the height at which profile is to be selected, removing curvature, smoothing, identifying groove locations are explained.

In the study conducted by \citet{aoas} a machine learning based algorithm was developed for same source matching of bullets and error rates were discussed using the database from the Hamby Study \citep{hamby}. In this paper, we first try to adapt a deterministic algorithm and method developed for toolmarks by \citet{chumbley} and improved by \citet{hadler}, to bullets. Then we consequently discuss about the efforts in doing so along with the associated error rates. The data used in this paper also belongs to the Hamby Study \citep{hamby}. This gives us a common platform for comparing the performance of the chumbley method on bullets with an already existing method proposed by \citet{aoas} for bullets. The proposed algorithm and method of \citet{chumbley}, in their paper, compares two toolmarks with the intention of determining if it comes from the same source (same tool). The method also provides a means to determine error rates and claims to reduce subject bias. As mentioned earlier, subject bias and error rate determination have been a long standing issue in firearm examination \citep{NAS:2009}. This remains one of the motivations to explore the adaptibility of the Chumbley score methodology to bullets. \citet{chumbley} used an empirical based setup to validate their proposed algorithm and quantitative method which calculates a U-statisic for the purpose of classification of toolmarks as matching or non-matching. The data for their study was obtained from 50 sequentially manufactured screwdriver tips, and preselected comparison window sizes were given as inputs to the algorithm. The algorithm then compares the two toolmarks and comes up with a U-statistic and an associated p-value to designate them as matches or non-matches. The performance for every 100 comparisons, of the algorithm proposed by \citet{chumbley} and the improvement proposed by \citet{hadler} are listed in the table below.

```{r, results='asis'}
library(knitr)
library(kableExtra)
library(xtable)

a1<- matrix(NA, ncol = 3, nrow = 2)
toolmark1.confusion<- data.frame(a1)
colnames(toolmark1.confusion)<- c("Classification", "Match", "Non-Match")
toolmark1.confusion$Classification<- c("Match", "Non-Match")
toolmark1.confusion$Match<- c(41,2)
toolmark1.confusion$`Non-Match`<- c(9,48)

a2<- matrix(NA, ncol = 3, nrow = 2)
toolmark2.confusion<- data.frame(a2)
colnames(toolmark2.confusion)<- c("Classification", "Match", "Non-Match")
toolmark2.confusion$Classification<- c("Match", "Non-Match")
toolmark2.confusion$Match<- c(47,0)
toolmark2.confusion$`Non-Match`<- c(3,50)

chumbley.confusion<-knitr::kable(toolmark1.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))
hadler.confusion<-knitr::kable(toolmark2.confusion, digits=0, format = "latex" ,booktabs=TRUE) #%>% kableExtra::kable_styling(latex_options = c("hold"))

cat(c("\\begin{table}[!htb]
    \\begin{minipage}{.5\\linewidth}
      \\caption{Chumbley et al. 2010}
      \\centering",
        chumbley.confusion,
    "\\end{minipage}%
    \\begin{minipage}{.5\\linewidth}
      \\centering
        \\caption{Hadler et. al. (2017)}",
       hadler.confusion,
    "\\end{minipage} 
\\end{table}"
))

```
<!-- Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citet{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citet{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citet{riva}, although as noted by \citet{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citet{efron}, but there are alternate error estimation techniques as described by \citet{aoas}, \citet{efron} and  \citet{vorburger2016} which give better results. -->

<!-- In case of methods that do not use machine learning algorithm, matching two striation marks with each other in order to identify if it is from the same source, can also be done using a well defined comparative statistical algorithm. The statistical algorithm is therefore expected  to go through a step by step procedure of identifying the class characteristics i.e striation marks for bullets, choosing the striae, and then follow a systematic procedure of comparing two striae with each other on the basis of criteria like cross-correlation (or others). Such a methodology would make it possible to determine error rates too and give definite results regarding what proportion of cases would the analysis be successful in identifying a match correctly and in how many cases would it end up being a false positive. -->

<!-- Identification criteria like consecutively matching striae (CMS) as first seen by Biasotti 1959 \citet{biasotti} and later mentioned in \citet{chu2013} more often than not include subject bias and are error prone and as seen in the work of \citet{miller} the number of CMS may turn out to be high even when the bullet is not fired by the same firearm. Parameters like cross-correlation factor as seen in the extensive comparisons made by \citep{aoas} seem to perform much better for matching purposes. -->

<!-- An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features. The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks. \citet{aoas} -->

<!-- The method employed by \citet{aoas} uses the cross correlation factor as a means to identify the stable region. The markings or striae in this region are supposed to have a high CCF with each other. Therefore a Bullet profile is chosen from this region by taking the cross section at a given height. A loess fit on the bullet profile produces  residuals which are termed as signatures and can be used as a unique identifier that can be used while matching two bullets to a firearm. -->

<!-- Compairing pairs of toolmarks on the otherhand, with the intention of matching it to a tool has been studied relatively more in the past as compared to bullets, and \citet{chumbley} have described in their paper an algorithm and analytic method that compares two toolmarks and come to the conclusion if they are from the same tool or not. The method also determines the error rates, reduces subject bias and designate the two toolmarks as matches or non-matches with respect to a source. \citet{chumbley} used an empirical based setup to validate their proposed analytic and quantitative algorithm. -->

<!-- \citet{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citet{hadler} on the other hand, in the improved version of the algorithm first described by \citet{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches. -->
<!-- ```{r} -->
<!-- a<- matrix(NA, ncol = 3, nrow = 2) -->
<!-- toolmark.falsepositive<- data.frame(a) -->
<!-- colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match") -->
<!-- toolmark.falsepositive$Classification<- c("Match", "Non-Match") -->
<!-- toolmark.falsepositive$Match<- c(47,0) -->
<!-- toolmark.falsepositive$`Non-Match`<- c(3,50) -->

<!-- knitr::kable(toolmark.falsepositive, digits=0, format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("hold")) -->
<!-- ``` -->

@@

```{r wo120-profile, fig.width = 6, fig.height=4, fig.cap="The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120"}
wo120_profile <- errors2 %>% filter(wo==120, coarse == 0.25) %>% 
  #gather(type, error, actual:beta) %>% 
  gather(type, error, beta) %>% 
  ggplot(aes(x = wv, y = error, colour = factor(alpha)))  + 
  geom_line() + geom_point() + facet_wrap(~type, scales="free") +
    theme_bw()+ 
  theme(legend.position="bottom") + ggtitle("Profile" , subtitle = "Varying validation window sizes, Optimization window = 120, Coarseness = 0.25")

@@

combined120
#wo120_profile
#wo120_1
#multiplot(wo120_1, wo120_profile1)
```
```{r failtest-coarseness, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}

miss_1
```
```{r failtest-wv, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}

miss_2
```
```{r failtest-wo, fig.width = 6, fig.height=4, fig.cap="The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120", eval= FALSE}

### 2018-02-05 doc/paper.tex
adding modified introduction

Striation marks are used to determine same source for bullets. Current
standards ask for a visual inspection of the marks under a comparison
microscope by a firearms examiner. One of the problems raised by the NAS
report in 2009 is the subjectivity involved in this comparison and the
need for determining error rates \citet{NAS:2009}. The issue of
subjectivity of a firearms examiner has been reviewed by many authors
where things like scientific principle testability and error rates are
considered to be defining aspects for an objective analysis such that
without identifying the uncertainities associated with a method of
comparison, it is inconclusive to regards an analysis complete. This
means that with the usual method of visual comparison that is adopted by
firearms examiners, lies the problem of lack of any systemized
quantification and as such there seems no certain way of associating it
with any kind of mathematical probability. Since determining error rates
has been duly noted as a fundamental problem in forensic science
\citet{NAS:2009}, there is a need of methods that address this problem.
Some of the methods that have been used to estimate error rates in
firearm examinations include Automatic cartridge case comparison where
same source identification of cartridge cases and associated error rates
have been provided by \citet{riva}, although as noted by \citet{aoas},
in this case alignments of striae involves roation of planes, which
cannot be generalized for bullets. In case of methods that use machine
learning algorithms some methods which are bootsrap based methods often
tend to give over-estimated error rates \citet{efron}, but there are
alternate error estimation techniques as described by \citet{aoas},
\citet{efron} and \citet{vorburger2016} which give better results.

In case of methods that do not use machine learning algorithm, matching
two striation marks with each other in order to identify if it is from
the same source, can also be done using a well defined comparative
statistical algorithm. The statistical algorithm is therefore expected
to go through a step by step procedure of identifying the class
characteristics i.e striation marks for bullets, choosing the striae,
and then follow a systematic procedure of comparing two striae with each
other on the basis of criteria like cross-correlation (or others). Such
a methodology would make it possible to determine error rates too and
give definite results regarding what proportion of cases would the
analysis be successful in identifying a match correctly and in how many
cases would it end up being a false positive.

Identification criteria like consecutively matching striae (CMS) as
first seen by Biasotti 1959 \citet{biasotti} and later mentioned in
\citet{chu2013} more often than not include subject bias and are error
prone and as seen in the work of \citet{miller} the number of CMS may
turn out to be high even when the bullet is not fired by the same
firearm. Parameters like cross-correlation factor as seen in the
extensive comparisons made by \citep{aoas} seem to perform much better
for matching purposes.

An important aspect of same source matching in firearms is
identification of bullet microtopographies that are unique to be chosen
for a match, Bullet profiles and signatures prove to be such features.
The choice of a bullet signature to uniquely define a bullet depends on
finding first, a stable region on a Bullet land which minimal noise and
many pronounced striation marks. \citet{aoas}

The method employed by \citet{aoas} uses the cross correlation factor as
a means to identify the stable region. The markings or striae in this
region are supposed to have a high CCF with each other. Therefore a
Bullet profile is chosen from this region by taking the cross section at
a given height. A loess fit on the bullet profile produces residuals
which are termed as signatures and can be used as a unique identifier
that can be used while matching two bullets to a firearm.

Compairing pairs of toolmarks on the otherhand, with the intention of
matching it to a tool has been studied relatively more in the past as
compared to bullets, and \citet{chumbley} have described in their paper
an algorithm and analytic method that compares two toolmarks and come to
the conclusion if they are from the same tool or not. The method also
determines the error rates, reduces subject bias and designate the two
toolmarks as matches or non-matches with respect to a source.
\citet{chumbley} used an empirical based setup to validate their
proposed analytic and quantitative algorithm.

\citet{chumbley} found that the algorithm gives false-positives in
slightly over 2 cases (or 2\%) of the time and about 9 false-negatives
(or 9\%) for every 100 comparisons. \citet{hadler} on the other hand, in
the improved version of the algorithm first described by
\citet{chumbley} found false-positives for 0 cases and 3 false-negatives
(or 6\%) for 50 comparisons of knwon matches.

\begin{table}[!h]
\centering
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 47 & 3\\
Non-Match & 0 & 50\\
\bottomrule
\end{tabular}
\end{table}

@@

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-coarseness-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-coarseness}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-wv-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wv}
\end{figure}\begin{figure}

{\centering \includegraphics[width=\textwidth]{figures/failtest-wo-1}

\caption{The figure on the left shows the actual Type I error while the figure on the right shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:failtest-wo}
\end{figure}

>>>>

Compairing pairs of toolmarks with the intention of matching it to a
tool has been studied many times in the past. Extensive examples in
literature for tools and toolmark research ranging from screwdrivers to
groove pliers to slip-joint pliers can be found in the work of
\citet{manytoolmarks1}, \citet{manytoolmarks2}, \citet{miller},
\cite{afte-chumbley} and many more. In comparison to this, same source
matching of bullets to firearms has not been examined as prominently as
that of toolmarks, with even less information available on validity of
methods and error rates associated with firearms examination. The
National Academy of Sciences in its report in 2009 \citep{NAS:2009}
discussed the need for determining error rates in methods proposed for
firearms examination. In the context of same source matching and error
rates determination, as seen in the case of most forensic applications,
the first step involves identification of unique features that are
characteristic of the object at hand. For the case of bullets and
firearms, striation marks on the surface of the bullet are considered to
be such markings that can be used in methods for same source matching.
These marks are often a product of rifling and impurities and defects
due to manufacturing in the barrel of the gun, which leads to engravings
on the bullet surface \citep{afte-article1992}. In current practice,
firearm examiners invariably make visual comparisons of bullet striae
and use visual assesment tools to dignify bullets as being matches and
non-matches. One way of accomplishing anykind of comparison between
bullets is to do a comparison between surface marking of two or more
bullet lands. Bullet Lands are considered to be areas between grooves
made by the rifling action of the barrel. These marking are considered
to be unique. The land engraved markings or sometimes termed as Bullet
profiles \citep{aoas} are striation marks made on Bullet lands and often
used for these land to land comparisons. Bullet Signatures is another
word used in literature as seen in the work of \citet{chu2013} and
\citet{aoas}. In our context bullet signatures refer to a processed
version of the raw land engraved markings or profiles. The generation of
bullet signatures involves first extraction of a bullet profile by
taking the cross-sectional of the surface at a given height and then
using loess fits to model the structure. The residuals of this fit are
called signatures, which are considered to be noise free and a good
reflection of the class charecteristics and unique features of a bullet.
A more detailed version of the extraction technique of signatures is
discussed by \citet{aoas}, where comprehensive details about the height
at which profile is to be selected, removing curvature, smoothing,
identifying groove locations are explained.

In the study conducted by \citet{aoas} a machine learning based
algorithm was developed for same source matching of bullets and error
rates were discussed using the database from the Hamby Study
\citep{hamby}. In this paper, we first try to adapt a deterministic
algorithm and method developed for toolmarks by \citet{chumbley} and
improved by \citet{hadler}, to bullets. Then we consequently discuss
about the efforts in doing so along with the associated error rates. The
data used in this paper also belongs to the Hamby Study \citep{hamby}.
This gives us a common platform for comparing the performance of the
chumbley method on bullets with an already existing method proposed by
\citet{aoas} for bullets. The proposed algorithm and method of
\citet{chumbley}, in their paper, compares two toolmarks with the
intention of determining if it comes from the same source (same tool).
The method also provides a means to determine error rates and claims to
reduce subject bias. As mentioned earlier, subject bias and error rate
determination have been a long standing issue in firearm examination
\citep{NAS:2009}. This remains one of the motivations to explore the
adaptibility of the Chumbley score methodology to bullets.
\citet{chumbley} used an empirical based setup to validate their
proposed algorithm and quantitative method which calculates a U-statisic
for the purpose of classification of toolmarks as matching or
non-matching. The data for their study was obtained from 50 sequentially
manufactured screwdriver tips, and preselected comparison window sizes
were given as inputs to the algorithm. The algorithm then compares the
two toolmarks and comes up with a U-statistic and an associated p-value
to designate them as matches or non-matches. The performance for every
100 comparisons, of the algorithm proposed by \citet{chumbley} and the
improvement proposed by \citet{hadler} are listed in the table below.

\begin{table}[!htb]
    \begin{minipage}{.5\linewidth}
      \caption{Chumbley et al. 2010}
      \centering 
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 41 & 9\\
Non-Match & 2 & 48\\
\bottomrule
\end{tabular} \end{minipage}%
    \begin{minipage}{.5\linewidth}
      \centering
        \caption{Hadler et. al. (2017)} 
\begin{tabular}{lrr}
\toprule
Classification & Match & Non-Match\\
\midrule
Match & 47 & 3\\
Non-Match & 0 & 50\\
\bottomrule
\end{tabular} \end{minipage} 
\end{table}

@@

\caption{The figure shows the Type II error for different validation window sizes and different chosen nominal alpha levels when the size of the optimization window = 120}\label{fig:wo120-profile}
\end{figure}
